### 2022-05-13 MSC-GUDMUNDSSON-2022.tex
Tiny fixes. Thesis in final mode. COMPLETE.

%%%%%%% CHOOSE ONE OF THESE %%%%%%%%%%%%%%%
%% projectreport: Project report (CS)
%% bachelors: Bachelor of Science thesis
%% masters: Master of Science thesis
%% doctorate: Doctor of Philosophy dissertation
%
%%%%%%% CHOOSE ONE OF THESE %%%%%%
%% 
%% draft: speed up processing by skipping graphics and adding useful
%%     information for editing.  Also sets spacing to double so that it is easier to
%%     write editing marks on paper copy.
%% proof:  proofreading version (final formatting with warnings)
%% final: generate document for submission, removing FIXMEs, and
%%     other markup.  Throw error if any fatal FIXMEs still in document.
%%
%%%%%%% CHOOSE ONE OF THESE IF APPLICABLE %%%%%%
%%
%% deptsse: School of Science and Engineering
%% deptscs: School of Computer Science
%%
%%%%%%%% CHOOSE ANY COMBINATION OF THESE %%%%%%%%%%%%
%%
%% forcegraphics: force graphics, etc. to be included, even in draft mode
%% debug:  writes more messages to the log file, adds debugging output 
%%     and sizing boxes
%% icelandic: thesis is in Icelandic
%% oldstyle:  use the PhD headers and footers from the old CS template
%% online: for online versions (skip blank pages)
\documentclass[online,masters,deptscs,forcegraphics,draft]{ruthesis}

@@

%% Date in english and icelandic
%% NOTE: THIS IS THE DATE OF THE SUPERVISOR'S SIGNATURE!!!!!!
%% Predefined: \MLjan, \MLfeb, \MLmar, ... \MLdec
%\whensigned{day}{month}{year} %day is only used on some formats, but you must put something.
\whensigned{10}{\MLmay}{2022}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%% Bachelors Only!!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setID{160992--2989}%kennitala
\setSemester{2022-2}
\setShortSignedDate{10.4.2022}

@@

}  
{
Heimurinn er að færast frá því að framleiða raforku með jarðefnaeldsneyti, í átt að umhverfisvænni leiðum til raforkuframleiðslu eins og sólarorku. Sólarorka er óútreiknanleg og mjög háð veðurfari, aðallega skýjahulu. Óstöðugleiki og óútreiknanleiki eru óæskilegir þegar kemur að sólarorku, þar sem framleiðsla verður að vera jöfn neyslu. Til að tryggja afhendingu raforku þegar sólarorka er notuð, reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði. Sólarorkuframleiðsla yrði skilvirkari og áreiðanlegri ef hægt væri að spá nákvæmar um framleiðslugetu dagsins, t.d. með áreiðanlegri veðurspám og áhrif þeirra á framleiðslugetu.
Verðmæt gögn eru til í formi veðurspáa sem hægt er að nýta til að gera sólarorkuframleiðslu útreiknanlegri. Við notum nýjar aðferðir á sviði djúpnáms, athygli(e. attention), og þá sérstaklega tímasamrunabylta (e. temporal fusion transformers) til að spá fyrir um sólargeislun við sólarorkuver sem fylgir náið raforkuframleiðslu orkuversins. Við byggjum þessa vinnu á háupplausnarveðurspám Belgings ehf., 24 klst. geislunarspám fyrir staðsetningu okuversins.
Módelið nær meðal meðaltölugildisprósentuskekkju (e. mean absolute percentage error) upp á 13,21\% á prófunargagnasettinu fyrir eins dags spár. Að auki skilar módelið öryggisbili fyrir spárnar, sem veitir raforkuframleiðendum innsæi um það hvort þeir geti búist við óútreiknanlegri sólarorkuframleiðslu. Módelið skilar nákvæmari spám þegar það styðst við veðurspágögn af stærra svæði í nánd orkuversins en þegar það nýtir eingöngu gögn fyrir nærumhverfi orkuversins.
} % Icelandic abstract goes here
\setAbstract{\MLAbstractText}

>>>>

%%%%%%% CHOOSE ONE OF THESE %%%%%%%%%%%%%%%
%% projectreport: Project report (CS)
%% bachelors: Bachelor of Science thesis
%% masters: Master of Science thesis
%% doctorate: Doctor of Philosophy dissertation
%
%%%%%%% CHOOSE ONE OF THESE %%%%%%
%% 
%% draft: speed up processing by skipping graphics and adding useful
%%     information for editing.  Also sets spacing to double so that it is easier to
%%     write editing marks on paper copy.
%% proof:  proofreading version (final formatting with warnings)
%% final: generate document for submission, removing FIXMEs, and
%%     other markup.  Throw error if any fatal FIXMEs still in document.
%%
%%%%%%% CHOOSE ONE OF THESE IF APPLICABLE %%%%%%
%%
%% deptsse: School of Science and Engineering
%% deptscs: School of Computer Science
%%
%%%%%%%% CHOOSE ANY COMBINATION OF THESE %%%%%%%%%%%%
%%
%% forcegraphics: force graphics, etc. to be included, even in draft mode
%% debug:  writes more messages to the log file, adds debugging output 
%%     and sizing boxes
%% icelandic: thesis is in Icelandic
%% oldstyle:  use the PhD headers and footers from the old CS template
%% online: for online versions (skip blank pages)
\documentclass[online,masters,deptscs,forcegraphics,final]{ruthesis}

@@

%% Date in english and icelandic
%% NOTE: THIS IS THE DATE OF THE SUPERVISOR'S SIGNATURE!!!!!!
%% Predefined: \MLjan, \MLfeb, \MLmar, ... \MLdec
%\whensigned{day}{month}{year} %day is only used on some formats, but you must put something.
\whensigned{20}{\MLmay}{2022}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%% Bachelors Only!!! %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setID{160992--2989}%kennitala
\setSemester{2022-2}
\setShortSignedDate{20.5.2022}

@@

}  
{
Heimurinn er að færast frá því að framleiða raforku með jarðefnaeldsneyti, í átt að umhverfisvænni leiðum til raforkuframleiðslu eins og sólarorku. Sólarorka er óútreiknanleg og mjög háð veðurfari, aðallega skýjahulu. Óstöðugleiki og óútreiknanleiki eru óæskilegir þegar kemur að sólarorku, þar sem framleiðsla verður að vera jöfn neyslu. Til að tryggja afhendingu raforku þegar sólarorka er notuð reiða framleiðendur sig oft á dýr og mengandi orkuver sem ganga fyrir jarðefnaeldsneyti og standa í viðbragðsstöðu til að bregðast við óvæntum breytingum á sólarorkuframboði. Sólarorkuframleiðsla yrði skilvirkari og áreiðanlegri ef hægt væri að spá nákvæmar um framleiðslugetu dagsins, t.d. með áreiðanlegri veðurspám og áhrif þeirra á framleiðslugetu.
Verðmæt gögn eru til í formi veðurspáa sem hægt er að nýta til að gera sólarorkuframleiðslu útreiknanlegri. Við notum nýjar aðferðir á sviði djúpnáms, athygli(e. attention), og þá sérstaklega tímasamrunabylta (e. temporal fusion transformers) til að spá fyrir um sólargeislun við sólarorkuver sem fylgir náið raforkuframleiðslu orkuversins. Við byggjum þessa vinnu á háupplausnarveðurspám Belgings ehf., 24 klst. geislunarspám fyrir staðsetningu okuversins.
Módelið nær meðal meðaltölugildisprósentuskekkju (e. mean absolute percentage error) upp á 13,21\% á prófunargagnasettinu fyrir eins dags spár. Að auki skilar módelið öryggisbili fyrir spárnar sem veitir raforkuframleiðendum innsæi um það hvort þeir geti búist við óútreiknanlegri sólarorkuframleiðslu. Módelið skilar nákvæmari spám þegar það styðst við veðurspágögn af stærra svæði í nánd orkuversins en þegar það nýtir eingöngu gögn fyrir nærumhverfi orkuversins.
} % Icelandic abstract goes here
\setAbstract{\MLAbstractText}

### 2022-05-13 methods.tex
Tiny fixes. Thesis in final mode. COMPLETE.

@@

    \afterpage{%
    \begin{figure}
        \centering
         \makebox[\textwidth][c]{\includegraphics[scale=0.5]{imgs/TFT_overview.jpg}}
        \caption{An overview of the Temporal Fusion Transformer.
        \label{fig:tft_overviwew}}
    \end{figure}
    \clearpage
    }

@@

    \clearpage
    \noindent\textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}. Some variables may be more important than others, \enquote{TFT is designed to provide instance-wise variable selection through the use of variable selection networks applied to both static covariates and time-dependent covariates.} \cite{lim_temporal_2020} The variable selection mechanism also allows the model to remove noisy inputs that might have a negative effect on performance. \\
    \textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
    \textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, display a confidence interval for the predictions \cite{lim_temporal_2020}. \\
    Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.

    \afterpage{%
    \begin{figure}
        \centering
         \makebox[\textwidth][c]{\includegraphics[scale=0.36]{imgs/TFT.png}}
        \caption{A detailed view of the architecture of the Temporal Fusion transformer. \cite{lim_temporal_2020}
        \label{fig:ttgt_detail}}
    \end{figure}
    \clearpage
    }

@@

    \afterpage{%
    \begin{figure}
        \centering
        \makebox[\textwidth][c]{\includegraphics[scale=0.6]{imgs/solar_TFT_overview.jpg}}
        \caption{An overview of the setup of the neural network.
        \label{fig:solar_tft_overviwew}}
    \end{figure}
    \clearpage
    }

>>>>

\newpage

@@

    \afterpage{%
    \begin{figure}
        \centering
         \makebox[\textwidth][c]{\includegraphics[scale=0.5]{imgs/TFT_overview.jpg}}
        \caption{An overview of the Temporal Fusion Transformer \cite{lim_temporal_2020}.
        \label{fig:tft_overviwew}}
    \end{figure}
    \clearpage
    }

@@

    \noindent\textbf{Variable selection networks} emphasize the most relevant inputs for every time step \cite{lim_temporal_2020}. Some variables may be more important than others, \enquote{TFT is designed to provide instance-wise variable selection through the use of variable selection networks applied to both static covariates and time-dependent covariates.} \cite{lim_temporal_2020} The variable selection mechanism also allows the model to remove noisy inputs that might have a negative effect on performance. \\
    \textbf{Static covariate encoders} encode static features into context vectors, integrating them into the network \cite{lim_temporal_2020}.  \\
    \textbf{Temporal processing} using a more traditional RNN sequence-to-sequence layer (see Section~\ref{cha:RNN}) for short term, local processing and the more novel multi-headed attention layer (see Section~\ref{cha:transformer}) for long term pattern capture \cite{lim_temporal_2020}.\\
    \textbf{Prediction intervals} in the form of quantile forecasts, discussed in Section~\ref{sec:loss_function}, display a confidence interval for the predictions \cite{lim_temporal_2020}. \\
    Figure~\ref{fig:ttgt_detail} shows a detailed view of the TFT structure as discussed.

    \afterpage{%
    \begin{figure}
        \centering
         \makebox[\textwidth][c]{\includegraphics[scale=0.3]{imgs/TFT.png}}
        \caption{A detailed view of the architecture of the Temporal Fusion transformer \cite{lim_temporal_2020}.
        \label{fig:ttgt_detail}}
    \end{figure}
    \clearpage
    }

@@

    \afterpage{%
    \begin{figure}
        \centering
        \makebox[\textwidth][c]{\includegraphics[scale=0.5]{imgs/solar_TFT_overview.jpg}}
        \caption{An overview of the setup of the neural network.
        \label{fig:solar_tft_overviwew}}
    \end{figure}
    \clearpage
    }
